### 2019-10-01 paper/latex/template.tex
Updates to paper and paperover code to export to PSB format

%\documentclass[wsdraft]{ws-procs11x85}

\documentclass{ws-procs11x85}
\usepackage{ws-procs-thm}           % comment this line when `amsthm / theorem / ntheorem` package is used
\usepackage{hyperref}
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

\begin{document}

\title{TITLE}

\author{Jake Lever\textsuperscript{1}, Julia M. Barbarino\textsuperscript{2}, Li Gong\textsuperscript{2}, Rachel Huddart\textsuperscript{2}, Katrin Sangkuhl\textsuperscript{2}, Ryan Whaley\textsuperscript{2}, Michelle Whirl-Carrillo\textsuperscript{2}, Mark Woon\textsuperscript{2}, Teri E. Klein\textsuperscript{2,3} and Russ B. Altman\textsuperscript{1-3}$^\dag$}

\address{
\textsuperscript{1}Department of Bioengineering, Stanford University, Stanford, CA, 94305 \\
\textsuperscript{2}Department of Biomedical Data Science, Stanford University, Stanford, CA, 94305 \\
\textsuperscript{3}Department of Medicine, Stanford University, Stanford, CA, 94305 \\
$^\dag$E-mail: rbaltman@stanford.edu\\
}

\begin{abstract}
ABSTRACT
\end{abstract}

\keywords{Pharmacogenomics; Biocuration; Text mining.}

% required, do-not-remove
\copyrightinfo{\copyright\ 2019 The Authors. Open Access chapter published by World Scientific Publishing Company and distributed under the terms of the Creative Commons Attribution Non-Commercial (CC BY-NC) 4.0 License.}

CONTENTS

\bibliographystyle{ws-procs11x85}
\bibliography{bibliography}

\end{document}

%%% \renewcommand\bibname{References\\ {\normalfont\it References can be typed in your preferred bibliography style.}}